---
title: "Practical Machine Learning Project"
author: "Xiaoya Zhu"
date: "04/25/2015"
output: html_document
---

## Synopsis

        This data analysis is to explore the HAR (Human Activity Recognition) Weight Lifting Exercises Dataset. The goal is to predict the manner in which they did the exercise. 

Read more information from this web site: 

[HAR HOME](http://groupware.les.inf.puc-rio.br/har)

## Data Processing 
####    Download file and load data.

The training data can be downloaded from this web site:
[training](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data can be downloaded from this web site:
[testing](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv).

```{r,downloadfile, message=F, warning=F}
library(caret)
library(rpart)
library(randomForest)
library(AppliedPredictiveModeling)

trainingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if(!file.exists("training.csv"))
download.file(trainingURL, destfile="training.csv", method="curl")
if(!file.exists("testing.csv"))
download.file(testingURL, destfile="testing.csv", method="curl")
training <- read.csv("training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("testing.csv", na.strings=c("NA","#DIV/0!",""))
dim(training)
dim(testing)
```

#### Clean data

Cast column from 8 to the last column to be numeric.
Remove column 1 to 7 which are not related with our model.
Remove columns only have value NA or 0.

```{r}
for(i in c(8:ncol(training)-1)) {training[,i] = as.numeric(as.character(training[,i]))}
for(i in c(8:ncol(testing)-1)) {testing[,i] = as.numeric(as.character(testing[,i]))}
columns <- colnames(training[colSums(is.na(training)) == 0])[-(1:7)]
training <- training[columns]

dim(training)
dim(testing)
```

#### Split training data set into 60% mytraining data set and 40% mytesting data set.
```{r}
idx <- createDataPartition(y=training$classe, p=0.6, list=FALSE )
mytraining <- training[idx,]
mytesting <-  training[-idx,]
dim(mytraining)
dim(mytesting)
```

#### Using Random Forests Model for Prediction.

```{r}
set.seed(333)
x <- mytraining[-ncol(mytraining)]
y <- mytraining$classe
modFit <- randomForest(x,y, data=mytraining)
```

## Conclusions  

Using mytesting data set to validate the prediction result. It shows Accuracy 0.99. 
```{r}
pred2 <- predict(modFit, mytesting, type = "class")
confusionMatrix(pred2,mytesting$classe)
```

## Prediction Assignment Submission

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
answers <-  predict(modFit, testing[columns[columns!='classe']], type = "class")
pml_write_files(answers)
```









